---
output:
  html_document: default
  pdf_document: default
---
```{r eval=TRUE, echo=FALSE}
# Load the libraries you will use for the mini-project report
library(dplyr)        # A package for data manipulation
library(sf)           # Simple feature for R
library(spdep)        # Functions and tests for evaluating spatial patterns 
library(tidyr)        # Tools to create tidy data
library(INLA)         # Integrated Nested Laplace Approximation package
library(ggplot2)      # A package for creating maps and graphs
library(viridis)      # A package providing color palettes 
library(patchwork)    # A package to compose plots

# For tables in RMarkdown
library(knitr)
library(kableExtra)
```

```{r}
#Import Data
```


```{r}
load("../DS6_CrimeUttarPradesh/CrimeUttarPradesh.RData")
```

```{r}
kable(
  data %>%
    group_by(year) %>%
    summarise(observed = sum(rape),expected = sum(e_rape),SMR = sum(rape) / sum(e_rape),.groups = "drop"),booktabs = TRUE,caption = "Number of rape") %>%
  kable_styling(
    bootstrap_options = "striped",full_width = FALSE,position = "center")
```

```{r}
# Create the dataset
rape_data <- data.frame(
  year = 2001:2014,
  observed = c(1956, 1413, 908, 1395, 1214, 1313, 1646, 1869, 1758, 1532, 2040, 1960, 3047, 3462),
  expected = c(1524.186, 1570.059, 1615.931, 1661.804, 1707.676, 1753.549, 1799.421, 1845.293, 1891.166, 1937.038, 1982.911, 2028.783, 2074.655, 2120.528)
)

# Reshape for plotting
rape_long <- rape_data %>%
  pivot_longer(cols = c("observed", "expected"), names_to = "type", values_to = "count")

ggplot(rape_long, aes(x = year, y = count, color = type)) +
  geom_line(size = 1.2) +
  geom_point(size = 2) +
  theme_minimal() +
  labs(title = "Observed vs Expected Rape (2001–2014)", y = "Number of Rape", x = "Year")
```

```{r}
rape_data$SMR <- rape_data$observed / rape_data$expected

ggplot(rape_data, aes(x = year, y = SMR)) +
  geom_line(color = "darkred", size = 1.2) +
  geom_point(size = 2) +
  geom_hline(yintercept = 1, linetype = "dashed", color = "gray") +
  theme_minimal() +
  labs(title = "Standardized Mortality Ratio (SMR) for Rape", y = "SMR", x = "Year")
```


```{r}
# 1. Load spatial and CSV data (if not already done)
Uttar_Pradesh <- st_as_sf(carto_up)

# 2. Calculate mean SMR per district
mean_smr <- data %>%
  group_by(ID_area) %>%
  summarise(mean_smr_rape = mean(smr_rape, na.rm = TRUE))

# 3. Merge with shapefile
Uttar_Pradesh <- Uttar_Pradesh %>%
  left_join(mean_smr, by = "ID_area")

# 4. Plot: SMR shading + District Numbers
ggplot() +
  geom_sf(data = Uttar_Pradesh, aes(fill = mean_smr_rape), color = "black", size = 0.3) +
  geom_sf_text(data = Uttar_Pradesh, aes(label = ID_area), size = 3, color = "black") +
  scale_fill_viridis_c(option = "plasma", na.value = "grey90", name = "Mean SMR") +
  labs(title = "Rape in Uttar Pradesh",
       subtitle = "District-wise Mean SMR (2001–2014)") +
  theme_minimal() +
  theme(axis.text = element_blank(),
        axis.title = element_blank(),
        panel.grid = element_blank(),
        plot.title = element_text(size = 14, face = "bold"),
        plot.subtitle = element_text(size = 12))
```

```{r}
district_table <- data %>%
  select(Identifier = ID_area, District = dist) %>%
  distinct() %>%
  arrange(Identifier)
# Split into 3 columns
n <- nrow(district_table)
split_n <- ceiling(n / 3)

col1 <- district_table[1:split_n, ]
col2 <- district_table[(split_n + 1):(2 * split_n), ]
col3 <- district_table[(2 * split_n + 1):n, ]

# Pad with empty strings
fill_row <- data.frame(Identifier = "", District = "")

while (nrow(col1) < split_n) col1 <- rbind(col1, fill_row)
while (nrow(col2) < split_n) col2 <- rbind(col2, fill_row)
while (nrow(col3) < split_n) col3 <- rbind(col3, fill_row)

# Combine into final table
table1 <- data.frame(
  Identifier1 = col1$Identifier,
  District1 = col1$District,
  Identifier2 = col2$Identifier,
  District2 = col2$District,
  Identifier3 = col3$Identifier,
  District3 = col3$District
)

# Print the formatted table
kable(
  table1,
  col.names = c("ID", "District", "ID", "District", "ID", "District"),
  caption = "Table 1. District identifiers for Uttar Pradesh"
)
```
Exploratory data analysis was conducted prior to statistical modeling to gain an initial understanding of the spatial and temporal patterns of rape incidents in Uttar Pradesh. By aggregating data across all districts over time, we plotted the yearly variation in observed and expected cases, along with standardized incidence ratios. To explore long-term spatial trends, we calculated the mean SMR for each district across the study period and visualized the spatial distribution in Figure 1. These plots are interpreted alongside the temporal trends to highlight regions with consistently elevated risk and periods of significant increase, which guided the development of subsequent spatio-temporal models.

# Spatio-temporal model (type I interaction)
$$

\begin{eqnarray}
O_{it} &\sim & \text{Poisson}(\rho_{it} E_{it}) \\
\log \rho_{it} &= & b_0 + b_i + \gamma_t +  \delta_{it} \\
\boldsymbol{b} &= & \frac{1}{\sqrt{\tau_b}}(\sqrt{1-\phi}\boldsymbol{v}_{*} + \sqrt{\phi}\boldsymbol{u}_{*})\\
\gamma_t & \sim & \hbox{RW(1)}\\
\delta_{it} & \sim & \hbox{Normal}(0, \sigma^2_{\delta})
\end{eqnarray}

$$
For area $i$ at time $t$,

-   $O$ stands for the number of observed cases , and $E$ is defined as the population scaled by 100,000,

-   $\rho$ is the incident rate of dengue fever for 100,000,

-   $Temp$ is the month maximum temperature, and $PDSI$ is the Palmer's drought severity index,

-   $\boldsymbol{b}$, the spatial random effect, is specified by a Besag-York-Molli'e (BYM) prior[@Besag1991] composed by $u_i$ and $v_i$, where $u_i$ is the spatially structured component defined by an intrinsic CAR prior [@Iarc2016]: $\boldsymbol{u}\sim ICAR(\boldsymbol{W}, \sigma^2_u)$, and $v_i$ is the unstructured component defined with prior: $v_s \overset{iid}{\sim} \text{Normal}(0,\sigma^2_v)$ ,

-   $\tau_b$ is the precision parameter controlling the marginal variance of the random effect with its prior defined as $P(\sigma_{\tau_b} > 0.5/0.31) = 0.01$, and $\phi$ is the mixing parameter measuring the proportion of the marginal variance with its prior defined as $P(\phi<0/5)=2/3$ [@Riebler2016].


```{r}
# Create neighbors list from sf object
Uttar_nb <- poly2nb(Uttar_Pradesh, snap = 1000, queen = TRUE)

# Step 2: Write INLA-compatible graph file
nb2INLA("Uttar.graph", Uttar_nb)
Uttar.adj = paste(getwd(),"/Uttar.graph",sep="")
```

```{r}
# Create space-time interaction index
data$ID.space.time = seq(1,dim(data)[1])

# INLA formula (type I interaction)
formula_ST_intI <- rape ~ 
  f(ID_area, model = "bym2", graph = Uttar.adj,
    hyper = list(
      prec = list(prior = "pc.prec", param = c(0.5 / 0.31, 0.01)),
      phi = list(prior = "pc", param = c(0.5, 2 / 3))
    )) +
  f(ID_year, model = "rw1", hyper = list(
      prec = list(prior = "pc.prec", param = c(0.5 / 0.31, 0.01))
    )) +
  f(ID.space.time, model = "iid", hyper = list(
      prec = list(prior = "pc.prec", param = c(0.5 / 0.31, 0.01))
    )) +
  offset(log(pop))

# Run model
stIntI.BYM.model <- inla(
  formula = formula_ST_intI,
  family = "poisson",
  data = data,
  control.compute = list(dic = TRUE, waic = TRUE)
)
```

```{r}
#Spatial Relative risks
RR_stIntI.BYM = c()

for(i in 1:70){
  RR_stIntI.BYM[i] = inla.emarginal(function(x) exp(x), 
        stIntI.BYM.model$marginals.random$ID.space[[i]])
}

#Posterior probabilities (for spatial RR)
RR_stIntI.BYM_marg = stIntI.BYM.model$marginals.random$ID.space[1:70]
PP_stIntI.BYM = lapply(RR_stIntI.BYM_marg, function(x) {1-inla.pmarginal(0,x)})

#Temporal Relative risks and CI95
RR_stIntI.RW_RR = c()
RR_stIntI.RW_lo = c()
RR_stIntI.RW_hi = c()

for(i in 1:14){
  #Posterior mean
  RR_stIntI.RW_RR[i] = inla.emarginal(function(x) exp(x), 
        stIntI.BYM.model$marginals.random$ID_year[[i]])
  #2.5% quantile 
  RR_stIntI.RW_lo[i] = inla.qmarginal(0.025,inla.tmarginal(function(x) exp(x), stIntI.BYM.model$marginals.random$ID_year[[i]]))
  #97.5% quantile 
  RR_stIntI.RW_hi[i] = inla.qmarginal(0.975, inla.tmarginal(function(x) exp(x), stIntI.BYM.model$marginals.random$ID_year[[i]]))
}

RR_stIntI.RW = data.frame(RR=RR_stIntI.RW_RR,low=RR_stIntI.RW_lo,high=RR_stIntI.RW_hi)

Temp2 = ggplot(RR_stIntI.RW, aes(seq(2001,2014), RR)) + geom_line() + ggtitle("ST model Int I") + geom_ribbon(aes(ymin=low,ymax=high), alpha=0.2) + labs(x="year")

Temp2
```

```{r}
# Create a dataframe with posterior RR and PP
resRR_PP_stIntI <- data.frame(
  resRR = RR_stIntI.BYM,
  PP = unlist(PP_stIntI.BYM),
  ID_area = unique(data$ID_area)  # Make sure there are 70 unique districts
)

# Categorize Relative Risk (RR) into intervals for mapping
resRR_PP_stIntI$resRRcat <- cut(
  resRR_PP_stIntI$resRR,
  breaks = c(
    min(resRR_PP_stIntI$resRR),
    0.4, 0.6, 0.8, 1.0, 1.2, 1.4, 1.6,
    max(resRR_PP_stIntI$resRR)
  ),
  include.lowest = TRUE
)

# Categorize Posterior Probability (PP) of RR > 1
resRR_PP_stIntI$PPcat <- cut(
  resRR_PP_stIntI$PP,
  breaks = c(0, 0.2, 0.8, 1.00),
  include.lowest = TRUE
)

# Merge with the Uttar Pradesh shapefile
Uttar_Pradesh <- left_join(Uttar_Pradesh, resRR_PP_stIntI, by = "ID_area")
```

```{r}
# Plot 1: Relative Risk categories
p5 <- ggplot() +
  geom_sf(data = Uttar_Pradesh, aes(fill = resRRcat), color = "black", size = 0.3) +
  theme_bw() +
  scale_fill_brewer(palette = "PuOr") +
  guides(fill = guide_legend(title = "RR")) +
  ggtitle("Relative Risk (RR)\nST Model Interaction I") +
  theme(
    text = element_text(size = 15),
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    plot.title = element_text(size = 12, face = "bold")
  )

# Plot 2: Posterior Probability categories
p6 <- ggplot() +
  geom_sf(data = Uttar_Pradesh, aes(fill = PPcat), color = "black", size = 0.3) +
  theme_bw() +
  scale_fill_viridis_d(
    option = "plasma",
    name = "P(RR > 1)",
    direction = -1,
    guide = guide_legend(
      title.position = 'top',
      reverse = TRUE
    )
  ) +
  ggtitle("Posterior Probability (PP)\nST Model Interaction I") +
  theme(
    text = element_text(size = 15),
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    plot.title = element_text(size = 12, face = "bold")
  )

# Combine the two maps side-by-side
(p5 | p6)
```

```{r}
# 1. Add the posterior mean of space-time interaction to your data
data$intI <- stIntI.BYM.model$summary.random$ID.space.time$mean

# 2. Categorize the interaction effect
data$intI_cat <- cut(data$intI, 
                     breaks = c(-1, -0.05, -0.01, 0.01, 0.05, 1), 
                     include.lowest = TRUE)

# 3. Merge with spatial map (Uttar_Pradesh) using ID_area
RESP_DATA_ST <- left_join(data, Uttar_Pradesh, by = "ID_area")

# 4. Convert to sf (if not already an sf object after join)
RESP_DATA_ST <- st_as_sf(RESP_DATA_ST)

# 5. Create year labels (2001–2014)
year_labels <- setNames(as.character(2001:2014), as.character(1:14))

# 6. Faceted plot of space-time interaction
ggplot() +
  geom_sf(data = RESP_DATA_ST, aes(fill = intI_cat), color = NA) +
  theme_bw() +
  scale_fill_brewer(palette = "PuOr", name = "Interaction") +
  guides(fill = guide_legend(title = NULL)) +
  theme(
    text = element_text(size = 16),
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    strip.text = element_text(size = 12)
  ) +
  facet_wrap(~ ID_year, ncol = , labeller = labeller(ID_year = year_labels)) +
  labs(title = "Space-Time Interaction Effect (ST Model Type I)")
```

```{r}
dat.hyper2 = 
  round(
  data.frame(median = stIntI.BYM.model$summary.hyperpar[,4],
    LL = stIntI.BYM.model$summary.hyperpar[,3], 
    UL = stIntI.BYM.model$summary.hyperpar[,5]),
  digits = 3)

row.names(dat.hyper2) = 
  rownames(stIntI.BYM.model$summary.hyperpar)

knitr::kable(dat.hyper2, caption = "Posterior median and 95% CrI of hyperparameters.") %>%  kable_styling(bootstrap_options = "striped", full_width = F, position = "center")
```

```{r}
dat.WAIC = data.frame(model = c("SpatTemp typeI"), 
                       WAIC = round(c(stIntI.BYM.model$waic$waic))
)

row.names(dat.WAIC) = NULL

knitr::kable(dat.WAIC, caption = "WAIC of the different models") %>%  kable_styling(bootstrap_options = "striped", full_width = F, position = "center")
```

