---
output:
  html_document: default
  pdf_document: default
---
```{r eval=TRUE, echo=FALSE,message=FALSE, warning=FALSE}
# Load the libraries you will use for the mini-project report
library(dplyr)        # A package for data manipulation
library(sf)           # Simple feature for R
library(spdep)        # Functions and tests for evaluating spatial patterns 
library(tidyr)        # Tools to create tidy data
library(INLA)         # Integrated Nested Laplace Approximation package
library(ggplot2)      # A package for creating maps and graphs
library(viridis)      # A package providing color palettes 
library(patchwork)    # A package to compose plots

# For tables in RMarkdown
library(knitr)
library(kableExtra)
library(sp)
```


```{r eval=TRUE, echo=FALSE,include=FALSE,message=FALSE, warning=FALSE}
load("/Users/zyy/Desktop/Bayesian project/CrimeUttarPradesh.RData")
carto_up <- st_as_sf(carto_up)
# 1 - Check the coordinates & min distances between polygons:

#st_coordinates(st_centroid(carto_up)) # these appear to be latitudes & longitudes
coord <- st_coordinates(st_centroid(carto_up))
dists <- spDists(coord)
#min(dists[dists > 0]) # min distance is 0.309

# 2- define neihbours list & convert to adjacency matrix

carto_up$reid <- carto_up$ID_area
carto_up_nb <- poly2nb(carto_up, snap=0.001, queen=TRUE) #  snap=0.01 seems the most accurate geographically
summary(carto_up_nb)
nb2INLA("map_adj",carto_up_nb) # create object with the location of the graph
adj = inla.read.graph(filename="map_adj") # store graph

data_st_rape = left_join(carto_up, data, by = "ID_area")
data_st_rape = data_st_rape %>% dplyr::rename(O=rape, E=e_rape)
data_st_dowry = left_join(carto_up, data, by = "ID_area")
data_st_dowry = data_st_dowry %>% dplyr::rename(O=dowry, E=e_dowry)
```



```{r eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE, results='hide'}
# Function to fit the model with custom PC priors
fit_st_model <- function(data, E, adj, u_val, alpha_val, phi_alpha) {
  formula_custom = O ~ 
    f(ID_area, model="bym2", graph=adj, 
      hyper=list(
        prec=list(prior="pc.prec", param=c(0.5/u_val, alpha_val)),
        phi=list(prior="pc", param=c(0.5, phi_alpha))
      )) + 
    f(ID_year, model="rw1", 
      hyper=list(prec=list(prior="pc.prec", param=c(0.5/u_val, alpha_val)))) + 
    f(ID_area_year, model="iid", 
      hyper=list(prec=list(prior="pc.prec", param=c(0.5/u_val, alpha_val))))
  
  model = inla(formula_custom, family="poisson", data=data, E=E, 
               control.compute = list(dic = TRUE, waic = TRUE),
               control.predictor = list(compute = TRUE))
  
  return(model)
}

# Define prior settings to test
# Define prior settings to test (updated names)
prior_settings = data.frame(
  model_id = c("Base_PC", "Weaker_Shrinkage", "Stronger_Spatial"),
  u_val = c(0.31, 0.5, 0.3),
  alpha_val = c(0.01, 0.01, 0.05),
  phi_alpha = c(2/3, 0.5, 0.8)
)

# Container for results
results = data.frame()

# Loop over models for Dowry and Rape
for(crime in c("dowry", "rape")) {
  data_input = if(crime == "dowry") data_st_dowry else data_st_rape
  
  for(i in 1:nrow(prior_settings)) {
    cat("Running", crime, "model:", prior_settings$model_id[i], "\n")
    
    model = fit_st_model(
      data = data_input,
      E = E,
      adj = adj,
      u_val = prior_settings$u_val[i],
      alpha_val = prior_settings$alpha_val[i],
      phi_alpha = prior_settings$phi_alpha[i]
    )
    
    # Extract hyperparameter summaries
    hyp = model$summary.hyperpar
    
    results = rbind(
      results,
      data.frame(
        Crime = crime,
        Model = prior_settings$model_id[i],
        DIC = model$dic$dic,
        WAIC = model$waic$waic,
        Prec_ID_area = hyp["Precision for ID_area", "mean"],
        Phi_ID_area = hyp["Phi for ID_area", "mean"],
        Prec_ID_year = hyp["Precision for ID_year", "mean"],
        Prec_ID_area_year = hyp["Precision for ID_area_year", "mean"]
      )
    )
  }
}
```

```{r}
kable(results, caption = "Sensitivity Analysis Results for Spatio-temporal Models with Interaction I")
```

To evaluate the robustness of the spatio-temporal models with Type I interaction to hyperprior assumptions, we conducted a sensitivity analysis using three different PC prior configurations for the precision parameters and the spatial mixing parameter ($\phi$) in the BYM2 model. The configurations included: 

1. the default prior (**Base_PC**) with $U = 0.31$, $\alpha = 0.01$, and $P(\phi < 0.5) = \frac{2}{3}$,  
2. a weakly informative prior (**Weaker_Shrinkage**) with $U = 0.5$, $\alpha = 0.01$, $P(\phi < 0.5) = 0.5$  
3. a stronger spatial structure prior (**Stronger_Spatial**) with $U = 0.3$, $\alpha = 0.05$, and $P(\phi < 0.5) = 0.8$. 

These settings were applied to both the dowry death and rape models, and model performance was assessed using the Deviance Information Criterion (DIC) and the Watanabe-Akaike Information Criterion (WAIC).

For the dowry death model, all three prior settings yielded nearly identical results ($\Delta$DIC < 0.15; $\Delta$WAIC < 0.25), with consistently high values of the spatial structure parameter $\phi$ (approximately 0.89), indicating strong spatial dependence and low sensitivity to prior specification. The rape model also showed minimal variation in DIC and WAIC across the prior settings, with slightly lower $\phi$ values (around 0.84–0.86), reflecting weaker spatial structure. 

Overall, the findings demonstrate that the spatio-temporal models are robust to reasonable variations in PC priors, and that the posterior inferences are primarily driven by the data rather than prior assumptions.

```{r}
knitr::opts_chunk$set(echo = TRUE)
# Create the table data
prior_settings <- data.frame(
  Model = c("Base_PC", "Weaker_Shrinkage", "Stronger_Spatial"),
  U = c(0.31, 0.5, 0.3),
  Alpha = c(0.01, 0.01, 0.05),
  Phi = c("2/3", "0.5", "0.8"),
  Description = c(
    "Original PC prior",
    "Weaker shrinkage: allows more variance, favors less spatial structure",
    "Stronger spatial structure: tighter σ, favors more spatial structure"
  )
)


# Render nicely with borders
kable(prior_settings, format = "html", escape = FALSE, align = "c",
      caption = "Hyperprior Settings for Sensitivity Analysis") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive", "bordered"),
                full_width = FALSE,
                position = "center") %>%
  column_spec(1:4, width = "8em") %>%
  column_spec(5, width = "30em")
```

