

# 1 - Set up

```{r}
# 1 - Check the coordinates & min distances between polygons:

#st_coordinates(st_centroid(carto_up)) # these appear to be latitudes & longitudes
coord <- st_coordinates(st_centroid(carto_up))
dists <- spDists(coord)
#min(dists[dists > 0]) # min distance is 0.309

# 2- define neihbours list & convert to adjacency matrix

carto_up$reid <- carto_up$ID_area
carto_up_nb <- poly2nb(carto_up, snap=0.001, queen=TRUE) #  snap=0.01 seems the most accurate geographically
summary(carto_up_nb)
nb2INLA("map_adj",carto_up_nb) # create object with the location of the graph
adj = inla.read.graph(filename="map_adj") # store graph


```

# 2 - spatial model
```{r}

# create dataset collaped across time
up_agg <- data %>% group_by(ID_area) %>% # aggregate over areas 
  summarise(O = sum(dowry), E = sum(e_dowry)) %>% # compute obs & expected
  mutate(SMR=O/E) # compute SMRs per area

# fit hierarchical poisson log-linear model
ID <- seq(1,70)
f_BYM2 <- O ~ f(ID, model="bym2", graph=adj,
                      hyper=list(
                        prec=list(prior = "pc.prec", param = c(0.5 / 0.31, 0.01)),
                        phi=list(prior="pc", param=c(0.5,2/3))
                      ))

s_model = inla(formula=f_BYM2, family="poisson", data=up_agg, E=E, 
               control.predictor = list(compute=TRUE), 
               control.compute = list(waic=TRUE))

```

Obtain posterior summary statistics

```{r}

# RRs
RR_s <- c()
for(i in 1:70) {
  RR_s[i] <- inla.emarginal(function(x) exp(x), s_model$marginals.random$ID[[i]])}

# Posterior probabilities

RR_s_marg <- s_model$marginals.random$ID[1:70]
PP_s <- lapply(RR_s_marg, function(x) {1-inla.pmarginal(0,x)})

s_RR_PP <- data.frame(resRR=RR_s, PP=unlist(PP_s), SP_ID=up_agg[,1])
```

Plot results

```{r, fig.width=12, fig.height=6}
breaks = c(min(s_RR_PP$resRR), 0.6, 0.8, 1, 1.2, 1.4, 1.6, 1.8, 2, max(s_RR_PP$resRR))
s_RR_PP$resRR_Cat = cut(s_RR_PP$resRR, breaks=breaks, include.lowest = TRUE)
breaks_pp = c(0, 0.2, 0.8, 1.00)
s_RR_PP$PP_cat = cut(s_RR_PP$PP, breaks_pp, include.lowest = TRUE)
map_s_RR_PP=left_join(carto_up, s_RR_PP, by=c("ID_area"="ID_area"))

map1_s = ggplot() + geom_sf(data=map_s_RR_PP) + aes(fill=resRR_Cat) + theme_bw() +
  scale_fill_brewer(palette = "PuOr") + guides(fill=guide_legend(title="RR")) + ggtitle("RR Spatial model") +
    theme(text = element_text(size=15), 
                  axis.text.x = element_blank(), 
                  axis.text.y = element_blank(), plot.title = element_text(size = 12, face = "bold"))

map2_s = ggplot() + geom_sf(data=map_s_RR_PP) + aes(fill=PP_cat) + theme_bw() +
    scale_fill_viridis(
    option = "plasma", name="PP",
    discrete = T,
    direction = -1,
    guide = guide_legend(
      title.position = 'top',
      reverse = T
    )) +  ggtitle("PP Spatial model") + theme(text = element_text(size=15), 
                  axis.text.x = element_blank(), 
                  axis.text.y = element_blank(), plot.title = element_text(size = 12, face = "bold")) 

map1_s | map2_s


```
Discussion: it seems that there is some clear spatial structure.

Get hyperparameter values

```{r}
s_model$summary.hyperpar
```
Discussion: majority (89%) of the spatial variability is explained by the spatially structured component, which is also supported by the visual assessment of the map.

# 3 - spatio-temporal model

```{r}
# Prep data 
data_st = left_join(carto_up, data, by = "ID_area")
data_st = data_st %>% dplyr::rename(O=dowry, E=e_dowry)

# Define & fit the model

formula_st = O ~ f(ID_area, model="bym2", graph=adj, hyper=list(
                        prec=list(prior = "pc.prec", param = c(0.5 / 0.31, 0.01)),
                        phi=list(prior="pc", param=c(0.5,2/3))
                      )) + 
  f(ID_year, model="rw1", hyper=list(prec=list(prior="pc.prec", param=c(0.5/0.31, 0.01))))

st_model = inla(formula=formula_st, family="poisson", data=data_st, E=E, control.compute=list(waic=TRUE), 
                control.predictor = list(compute=TRUE))

```

Plot the temporal residual RRs

```{r}
# derive posterior temporal RRs & CIs
RR_stRW_RR = c()
RR_stRW_l = c()
RR_stRW_h = c()

for(i in 1:14) {
  RR_stRW_RR[i] = inla.emarginal(function(x) exp(x), st_model$marginals.random$ID_year[[i]])
  RR_stRW_l[i] = inla.qmarginal(0.025,inla.tmarginal(function(x) exp(x), st_model$marginals.random$ID_year[[i]]))
  RR_stRW_h[i] = inla.qmarginal(0.975,inla.tmarginal(function(x) exp(x), st_model$marginals.random$ID_year[[i]]))
}

RR_stRW = data.frame(RR=RR_stRW_RR, low_CI=RR_stRW_l, up_CI = RR_stRW_h)

# Plot temporal residual RRs

temp1 = ggplot(RR_stRW, aes(seq(2001,2014), RR)) + geom_line() + ggtitle("ST model (no interactions)") + theme_bw() +
  geom_ribbon(aes(ymin=low_CI,ymax=up_CI), alpha=0.2) + labs(x="year")

temp1
```

map the spatial residual RRs & exceedance probabilities

```{r, fig.width=12, fig.height=6}
# derive spatial posterior means (RR) & exceedance probabilities

RR_stBYM = c()
for(i in 1:70) {RR_stBYM[i] = inla.emarginal(function(x) exp(x), st_model$marginals.random$ID_area[[i]])} ##
RR_stBYM_marg = st_model$marginals.random$ID_area[1:70]
PP_stBYM = lapply(RR_stBYM_marg, function(x) {1-inla.pmarginal(0,x)})

resRR_PP_st = data.frame(RR=RR_stBYM, PP=unlist(PP_stBYM), ID_area=carto_up$ID_area)

# plot 
breaks = c(min(resRR_PP_st$RR), 0.6, 0.8, 1, 1.2, 1.4, 1.6, 1.8, 2, max(resRR_PP_st$RR))
resRR_PP_st$RR_cat = cut(resRR_PP_st$RR, breaks=breaks, include.lowest = TRUE)
breaks_pp = c(0, 0.2, 0.8, 1.00)
resRR_PP_st$PP_cat = cut(resRR_PP_st$PP, breaks_pp, include.lowest = TRUE)
map_st_RR_PP=left_join(carto_up, resRR_PP_st, by=c("ID_area"="ID_area"))

map1_st = ggplot() + geom_sf(data=map_st_RR_PP) + aes(fill=RR_cat) + theme_bw() +
  scale_fill_brewer(palette = "PuOr") + guides(fill=guide_legend(title="RR")) + ggtitle("RR Spatio-temporal model") +
    theme(text = element_text(size=15), 
                  axis.text.x = element_blank(), 
                  axis.text.y = element_blank(), plot.title = element_text(size = 12, face = "bold"))

map2_st = ggplot() + geom_sf(data=map_st_RR_PP) + aes(fill=PP_cat) + theme_bw() +
    scale_fill_viridis(
    option = "plasma", name="PP",
    discrete = T,
    direction = -1,
    guide = guide_legend(
      title.position = 'top',
      reverse = T
    )) +  ggtitle("PP Spatio-temporal model") + theme(text = element_text(size=15), 
                  axis.text.x = element_blank(), 
                  axis.text.y = element_blank(), plot.title = element_text(size = 12, face = "bold")) 

(map1_s | map2_s) / (map1_st | map2_st)


```
Comment: this is identical to the spatial model? But I could not find any issue that would result in this.

# 4 Spatio-temporal model for type I interaction

```{r}
formula_st_int = O ~ f(ID_area, model="bym2", graph=adj, hyper=list(
                        prec=list(prior = "pc.prec", param = c(0.5 / 0.31, 0.01)),
                        phi=list(prior="pc", param=c(0.5,2/3))
                      )) + 
  f(ID_year, model="rw1", hyper=list(prec=list(prior="pc.prec", param=c(0.5/0.31, 0.01)))) + 
  f(ID_area_year, model="iid", hyper=list(prec=list(prior="pc.prec", param = c(0.5/0.31, 0.01))))

st_int_model = inla(formula = formula_st_int, family="poisson", data=data_st, E=E, 
                    control.compute=list(dic=TRUE, waic=TRUE))
```

```{r}
# temporal RRs & CIs

RR_st_int_RW_RR = c()
RR_st_intRW_l = c()
RR_st_int_RW_h = c()

for(i in 1:14) {
  RR_st_int_RW_RR[i] = inla.emarginal(function(x) exp(x), st_int_model$marginals.random$ID_year[[i]])
  RR_st_intRW_l[i] = inla.qmarginal(0.025,inla.tmarginal(function(x) exp(x), st_int_model$marginals.random$ID_year[[i]]))
  RR_st_int_RW_h[i] = inla.qmarginal(0.975,inla.tmarginal(function(x) exp(x), st_int_model$marginals.random$ID_year[[i]]))
}

RR_st_int_RW = data.frame(RR=RR_st_int_RW_RR, low_CI=RR_st_intRW_l, up_CI = RR_st_int_RW_h)

# Plot temporal residual RRs

temp2 = ggplot(RR_st_int_RW, aes(seq(2001,2014), RR)) + geom_line() + 
  ggtitle("ST model (Interaction I)") + theme_bw() +
  geom_ribbon(aes(ymin=low_CI,ymax=up_CI), alpha=0.2) + labs(x="year")

temp1 | temp2

```

```{r, fig.width=12, fig.height=12}
# spatial RRs

RR_st_int_BYM = c()
for(i in 1:70) {RR_st_int_BYM[i] = inla.emarginal(function(x) exp(x), 
                                                  st_int_model$marginals.random$ID_area[[i]])} ##
RR_st_intBYM_marg = st_int_model$marginals.random$ID_area[1:70]
PP_st_int_BYM = lapply(RR_st_intBYM_marg, function(x) {1-inla.pmarginal(0,x)})

# paste

resRR_PP_st_int = data.frame(RR=RR_st_int_BYM, PP=unlist(PP_st_int_BYM), ID_area=carto_up$ID_area)

# plot 
breaks = c(min(resRR_PP_st_int$RR), 0.6, 0.8, 1, 1.2, 1.4, 1.6, 1.8, 2, max(resRR_PP_st_int$RR))
resRR_PP_st_int$RR_cat = cut(resRR_PP_st_int$RR, breaks=breaks, include.lowest = TRUE)
breaks_pp = c(0, 0.2, 0.8, 1.00)
resRR_PP_st_int$PP_cat = cut(resRR_PP_st_int$PP, breaks_pp, include.lowest = TRUE)
map_st_int_RR_PP=left_join(carto_up, resRR_PP_st_int, by=c("ID_area"="ID_area"))

map1_st_int = ggplot() + geom_sf(data=map_st_int_RR_PP) + aes(fill=RR_cat) + theme_bw() +
  scale_fill_brewer(palette = "PuOr") + guides(fill=guide_legend(title="RR")) + ggtitle("RR Spatio-temporal model with Interaction I") +
    theme(text = element_text(size=15), 
                  axis.text.x = element_blank(), 
                  axis.text.y = element_blank(), plot.title = element_text(size = 12, face = "bold"))

map2_st_int = ggplot() + geom_sf(data=map_st_int_RR_PP) + aes(fill=PP_cat) + theme_bw() +
    scale_fill_viridis(
    option = "plasma", name="PP",
    discrete = T,
    direction = -1,
    guide = guide_legend(
      title.position = 'top',
      reverse = T
    )) +  ggtitle("PP Spatio-temporal model with Interaction I") + theme(text = element_text(size=15), 
                  axis.text.x = element_blank(), 
                  axis.text.y = element_blank(), plot.title = element_text(size = 12, face = "bold")) 

(map1_s | map2_s) / (map1_st | map2_st) / (map1_st_int | map2_st_int)
```
Comment: it seems like there is very little difference in spatial residuals. Some additional regions are highlighted to have a higher RR, and an extra region is highlighted to have a higher exceedance probability. 

Now plotting space-time interaction

```{r}
data_st$int = st_int_model$summary.random$ID_area_year$mean
data_st$int_cat = cut(data_st$int,  breaks=c(-1,-0.05, 
                  -0.01, 0.01, 0.05, 1),include.lowest = T)

ggplot() + geom_sf(data = data_st, aes(fill=int_cat)) + theme_bw() +
  scale_fill_brewer(palette = "PuOr") +
  guides(fill=guide_legend(title=NULL)) + 
  theme(text = element_text(size=20),
        axis.text.x = element_blank(), axis.text.y = element_blank()) +
facet_wrap(~ year, ncol = 4, labeller=labeller(ID_year=c("1"="2001","2"="2002","3"="2003","4"="2004","5"="2005",
                            "6"="2006","7"="2007","8"="2008","9"="2009","10"="2010",
                            "11"="2011","12"="2012","13"="2013", "14"="2014"))) +
labs("")
```

