---
title: "Bayesian Spatio-Temporal modelling of crime against women in Uttar Pradesh"
author: "Wanqing Fang, Ben Ling, Maria Serova, Yi Zhou"
date: "31/03/2025"
output:
  pdf_document: default
  html_document:
    fig_caption: true
    md_extensions: +raw_tex
  word_document: default
always_allow_html: true
header-includes: \usepackage{bm} \usepackage{booktabs}
bibliography: "biblio.bib"
abstract: "This project investigated spatio-temporal patterns of crimes against women, specifically rape and dowry deaths, in Uttar Pradesh, India, from 2001 to 2014. Bayesian hierarchical models (BHMs) using Besag–York–Mollié 2 (BYM2) priors and Type I interactions were applied to district-level data from the National Crime Records Bureau. Integrated Nested Laplace Approximation (INLA) facilitated efficient model fitting. Results confirmed distinct spatial patterns: rape risk was highest in the west and south around Aligarh and Chitrakoot respectively, while dowry deaths clustered in the southwest around Etah and Mainpuri. Temporally, rape incidents notably increased post-2012, possibly influenced by enhanced reporting following the Delhi gang rape incident, whereas dowry death rates remained stable. Incorporating spatio-temporal interactions improved model performance, capturing localised district-specific trends. These differences reveal latent underpinning factors which may be socio-cultural for example, suggesting targeted anti-dowry measures in hotspots and broader rape interventions statewide. Sensitivity analyses indicated that spatially structured effects dominated both crime types, with very little change seen by modifying prior assumptions. Methodologically, we validated  research by Vicente et al. (2018 and 2020), highlighting the importance of Bayesian spatio-temporal analyses and discuss future research to facilitate policy improvements, effective resource allocation and interventions addressing crimes against women."
editor_options:
  markdown:
    wrap: sentence
---

\pagenumbering{gobble} 
\pagenumbering{arabic}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE, fig.align = "center", class.source='klippy')
```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(position = c('top', 'right'),color = 'darkred',
               tooltip_message = 'Click to copy', tooltip_success = 'Done')
```

```{r eval=TRUE, echo=FALSE}
# Load the libraries 
library(dplyr)        
library(sf)           
library(spdep)        
library(tidyr)        
library(INLA)         
library(ggplot2)      
library(viridis)    
library(patchwork)    
library(knitr)
library(kableExtra)

# Prepare directory and data
#rm(list=ls())
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
load("CrimeUttarPradesh.RData")

# load figures
load("Figures.RData")
```

# Introduction

Violence against women remains a major public health and human rights issue in India.
Rape and dowry deaths are among the gravest forms, reflecting deep-rooted gender inequality.
Rape in particular is underreported due to social stigma and fear of retaliation [@Shanmugam2013].
The dowry system, whereby a bride’s family provides payment to the groom’s family, can lead to dowry deaths when brides are murdered or driven to suicide through harassment and violence [@Rastogi2006].
We look to Uttar Pradesh, India's most populous state as a reference point for analysis.

Vicente et al (2018) [@Vicente2018rape] first applied Bayesian Hierarchical Models (BHM) to the rape data, hypothesising a temporal effect from the widely publicised 2012 Delhi gang rape incident [@Delhigang].
They then expanded their methods to dowry deaths [@Vicente2020dowry].
Here, we build on that work by adopting Besag–York–Mollié 2 (BYM2) priors [@Besag1991], partitioning spatially structured and unstructured variation to validate and extend Vicente et al.’s conclusions [@Vicente2018rape].
Employing BHMs allows us to explicitly model spatial and temporal correlation and obtain more precise risk estimates via "strength borrowing" across spatial and temporal structure.
This approach is particularly valuable from a policy perspective, as it produces reliable district-level estimates even where crime reporting is sparse, facilitating more accurate resource allocation and potential targeted interventions.

Our spatiotemporal analyses of both crimes focus on the following questions:\
1.
Can we replicate Vicente et al.'s findings of distinct spatial patterns for rape and dowry deaths using BYM2 and Type I interactions?\
2.
How have the temporal trends for rape and dowry deaths evolved over the 14-year period, and what insights arise when examining the patterns of the two crimes in parallel?\
3.
Is there evidence of space-time interaction in the occurrence of these crimes, suggesting localised as opposed to state-wide spreading patterns?

By leveraging BHM techniques in this manner, we aim to produce more robust spatiotemporal insights that can guide targeted interventions and strengthen the literature on gender-based violence in India.

# Methods

Data on rape and dowry deaths were sourced from the National Crime Records Bureau, covering 70 districts in Uttar Pradesh, North India, from 2001 to 2014 [@NCRB2017].
Standardised morbidity ratios (SMRs) for rape and dowry deaths were provided in table form.
A consistency check confirmed that the expected values in SMR calculations were correctly derived from aggregated statewide data, ensuring a clear internal baseline for comparisons across districts.
Unsmoothed SMR data by district and year were mapped and graphed to visualise 'raw' baseline spatial and temporal patterns (Figure 1) as a starting point for our analysis.

BHMs were then employed to explicitly model spatial and temporal autocorrelation.
Each model was fitted separately for each crime, using Integrated Nested Laplace Approximation (INLA) to leverage its computational simplifications [@INLA09].

Three models were fitted: spatial (S), spatio-temporal (ST), and spatio-temporal with Interaction Type I (ST+I), specified as follows:

\normalsize
\renewcommand{\arraystretch}{1.0}
\setlength{\tabcolsep}{6pt}
\begin{center}
\begin{tabular}{p{0.33\textwidth} p{0.33\textwidth} p{0.33\textwidth}}

\textbf{S Model} &
\textbf{ST Model} &
\textbf{ST+I Model} \\

$O_i \sim \text{Poisson}(\rho_i E_i)$ &
$O_{it} \sim \text{Poisson}(\rho_{it} E_{it})$ &
$O_{it} \sim \text{Poisson}(\rho_{it} E_{it})$ \\

$\log \rho_i = b_0 + b_i$ &
$\log \rho_{it} = b_0 + b_i + \gamma_t$ &
$\log \rho_{it} = b_0 + b_i + \gamma_t + \delta_{it}$ \\

$\mathbf{b} = \dfrac{1}{\sqrt{\tau_b}} \left( \sqrt{1 - \phi} \mathbf{v_*} + \sqrt{\phi} \mathbf{u_*} \right)$ &
$\mathbf{b} = \dfrac{1}{\sqrt{\tau_b}} \left( \sqrt{1 - \phi} \mathbf{v_*} + \sqrt{\phi} \mathbf{u_*} \right)$ &
$\mathbf{b} = \dfrac{1}{\sqrt{\tau_b}} \left( \sqrt{1 - \phi} \mathbf{v_*} + \sqrt{\phi} \mathbf{u_*} \right)$ \\

&
$\gamma_t \sim \text{RW(1)}$ &
$\gamma_t \sim \text{RW(1)}$ \\

& &
$\delta_{it} \sim \mathcal{N}(0, \sigma^2_\delta)$ \\

\end{tabular}
\end{center}

Where, for area $i$ at time $t$, $O$ and $E$ denote the number of observed and expected cases, respectively, $\rho$ is the relative risk of crime, $b_0$ is the intercept term representing baseline log-risk.
$b_i$ is the spatial effect under BYM2 specification, decomposed into structured ($u_i$) and unstructured ($v_i$) components with a mixing parameter $\phi$ and overall precision $\tau_b$.
$\gamma_t$ denotes the temporal component modelled as a first-random walk ($RW1$), and $\sigma^2$ corresponds to a Type I interaction.

The BYM2 model was chosen for its decomposition of spatial variability into structured and unstructured components, which are easily controlled through the spatial mixing parameter $\phi$.
This offers clearer interpretability and more flexible spatial smoothing compared to standard BYM or intrinsic conditional autoregressive (ICAR) models, which either lack explicit decomposition (ICAR) or provide limited control over component mixing (standard BYM).

Penalised Complexity (PC) priors were used for all precision and spatial mixing parameters to impose regularisation and avoid overfitting.
All precision priors were defined as $P(\sigma > 0.5) = 0.01$, reflecting a strong prior belief in smooth effects across space and time, motivated by our assumption that risk changes gradually due to continuity of sociocultural and administrative characteristics.

The PC prior for spatial mixing parameter $\phi$, we defined as $P(\phi < 0.5) = \frac{2}{3}$ [@Riebler2016], reflecting our preference towards unstructured spatial variation.
This is driven by the assumption that district-specific factors (such as local crime reporting practices, governance effectiveness, opportunistic crime, or social norms) exert greater influence on crime risk than broader structured spatial trends like systemic poverty and organised crime.

To assess the robustness of our hyperprior choices, we conducted sensitivity analyses by implementing three Penalised Complexity (PC) prior configurations: Base_PC, Weaker_Shrinkage, and Stronger_Spatial, each varying in shrinkage strength and spatial structure preference, and compared model performance (WAIC) and hyperparameter summaries for both dowry death and rape models with Type I spatio-temporal interaction (Supplementary Table 3).

All ST and ST+I models were assessed using the Watanabe-Akaike Information Criterion (WAIC), chosen for its ability to balance model fit with effective penalisation of model complexity in BHMs.
The S model was excluded from WAIC comparison due to being fit on aggregate data.

# Results

Between 2001 and 2014, a total of 25,513 rape cases (mean: 2,043 ± 334 per year) and 28,600 dowry deaths (mean: 1,822 ± 691 per year) were reported.
Figure 1 shows the unsmoothed temporal trends and spatial variability of both crimes’ SMRs.

All three models identified comparable patterns of posterior spatial RRs (Supplementary Figures 2-3).
The ST+I model showed a better fit compared to the ST model, as reflected in its lower WAIC for both crimes (6240 vs 7781 for rape; 6318 vs 6654 for dowry deaths).
Its results are discussed further.

Sensitivity analyses using three PC prior configurations showed minimal variation in WAIC values, with consistently high spatial structure ($\phi$) estimates (\$\approx\$0.89 for dowry, \$\approx\$0.84–0.86 for rape), indicating that posterior inferences were primarily driven by the data rather than prior assumptions.(Figure 2)

Figure 2 shows the smoothed spatial posterior relative risks (RRs), as well as the posterior probabilities of exceeded risk (PPs).
Both unsmoothed and smoothed estimates of dowry deaths maintained an east-to-west gradient, with high-risk districts identified around Etah and Mainpuri in the southwest.
For rape, two high-risk clusters emerged around Aligarh in the west and Chitrakoot in the south.
Smoothed RRs ranged wider for rape (0.3–2.3) than dowry deaths (0.4–2.1), corroborated by the lower precision parameter of the spatial component for rape (Table 1).

Figure 3 presents temporal trends estimated by the ST and ST+I models, both of which are similar to the unsmoothed RRs (Figure 1).
The wider credible intervals in the ST+I model account for the district-specific variation captured by with the interaction term.

Spatio-temporal interactions indicated district-specific temporal deviations, with clearer clustering noted for dowry deaths than for rape (available in Supplementary Figure 4).
Drawing on these results, Figure 4 shows the posterior mean RRs of the ST+I model, combining spatial, temporal, and spatio-temporal components, which capture geographical and temporal dynamics.
While dowry death risks remained relatively stable after 2004, rape risks experienced a rapid resurgence after 2012, with several districts exhibiting RRs of 2.5–4.75 in 2014.

```{r EDA-figure, fig.width=12, fig.height=3, fig.cap="Figure 1: Temporal evolution of standardised SMRs for rape incidents and dowry deaths in Uttar Pradesh (2001-2014) (left) and spatial distribution of SMRs aggregated over the study period (right)"}

# Plot unsmoothed data
unsmoothed
```

```{r stInt_m, fig.width=10, fig.height=6, fig.cap="Figure 2: Spatial patterns of posterior mean relative risks (RR) and posterior probabilities (PP) exceeding 1 from spatio-temporal models with Type I interaction for rape incidents (top) and dowry deaths (bottom).", out.width="60%"}
stInt_m
```

```{r stInt_hyper_table, results='asis'}

stInt_hyper
```

```{r temp_st_stInt, fig.width=12, fig.height=3, fig.cap="Figure 3: Comparison of temporal trends in relative risks between spatio-temporal models without interaction (left) and with Type I interaction (right) for rape incidents and dowry deaths in Uttar Pradesh (2001-2014)."}
temp_st_stInt
```

```{r meanRR_Post, fig.width=8, fig.height=10, fig.cap="Figure 4: Annual posterior mean relative risks for rape incidents (top) and dowry deaths (bottom) across Uttar Pradesh districts from 2001 to 2014, demonstrating the evolution of spatial patterns over time.", out.width="60%"}
meanRR_post
```

# Discussion

The most notable finding is the distinct temporal pattern observed between the two crimes: rape incidence showed a substantial increase especially after 2012, while dowry deaths remained relatively stable (Figure 3).\
Vicente et al. (2018) reported similar findings for rape, noting its temporal alignment with the highly publicised 2012 Delhi gang rape incident and subsequent 2013 legal reforms that expanded the definition of rape [@Vicente2018rape] [@Delhigang].
The rape models also exhibited a markedly lower precision hyperparameter of temporal component, compared to that of dowry deaths (Table 1), suggesting that global temporal effects played a more prominent role in shaping rape risk compared to dowry risk.
Whilst our study cannot establish causation, the divergence in temporal trajectories of risk concurs that a sudden event affected rape incidence more than dowry deaths.

The high values of spatial mixing hyperparameter in both models (Table 1) point to the dominant contributions of structured spatial effects, consistent with observed spatial patterns.
Rape reporting showed a clear west-to-east gradient which was more marked in the later years (Figure 4), with higher rates in the western districts neighbouring Delhi.
This suggests that proximity to Delhi may have influenced reporting patterns, possibly through more effective awareness campaigns or stronger identification with the Delhi incident in western regions.

In contrast, the persistent clustering of dowry deaths around nearby southwestern districts like Etah and Mainpuri did not change with the 2012 incident.
It is noted that marital rape was not included in the widening of the definition of rape in early 2013 [@Lawchange].
While merely speculative, this potentially limited reporting by women experiencing chronic domestic abuse that may have ultimately resulted in dowry deaths.
The more consistent spatial clustering observed for dowry deaths compared to rape incidents likely reflects different underlying socio-cultural dynamics.
Dowry deaths may represent the culmination of prolonged domestic violence patterns tied to district-specific cultural factors, resulting in less heterogeneity compared to rape which may be more 'acute' and thus random in comparison.

We were not able to explain why both crimes showed an initial decline between 2001 to 2003, but infer that there were common statewide, or even national factors affecting both crimes, in contrast to the post-2012 period.

The superior performance of the STI+I model highlights the importance of at least checking for localised temporal trajectories in crime data analysis.
The assumption that all districts follow parallel temporal trends, implicit in non-interaction models (ST), may not capture the complex reality of violent crime.

From a policy perspective, our findings suggest different intervention approaches may be needed: targeted enforcement of anti-dowry laws in *specific* districts, versus broader educational and legal campaigns addressing rape *across the state* given its increased variability.
The district-specific deviations identified for rape could *then* inform to what degree localised prevention strategies should build on statewide efforts.
Further research into the spatial Delhi 'spillover' effect may bring interesting insights in helping to develop these interventions.

Limitations include the absence of district-level covariates constraining interpretation and the study period ending in 2014.
Future research should incorporate socioeconomic variables, for example internet coverage, explore different interaction models (II-IV), and conduct qualitative studies to better understand district-level variations in crime reporting and responses.

In conclusion, this spatio-temporal analysis demonstrates the value of advanced Bayesian modelling techniques for understanding the complex dynamics of gender-based violence.
We validated Vicente et al.'s findings of rape and dowry deaths' spatiotemporal patterns and demonstrated that they were robust to hyperprior variation [@Vicente2018rape] [@Vicente2020dowry].
We detailed the temporal trends of both crimes and showed the value of adding an interaction term in unveiling how different districts varied individually with time, while taking smoothing into account.
The distinct spatial patterns, temporal trends, and interaction effects observed for rape and dowry deaths underscore the importance of these models to aid crime-specific, geographically targeted interventions to address violence against women in Uttar Pradesh.

==============================

# Supplementary material

```{r dist_table, fig.width=12, fig.height=12, fig.cap="Supplementary Table 1"}
reg_table
```

```{r setting_table, fig.width=12, fig.height=12, fig.cap="Supplementary Table 2"}
sensitivity_priors
```

```{r sensitive_table, fig.width=12, fig.height=12, fig.cap="Supplementary Table 3"}
sensitivity
```

```{r reg_temp_smr, fig.width=12, fig.height=3, fig.cap="Supplementary Figure 1: Crude rates of rape incidents and dowry deaths in Uttar Pradesh (2001-2014), showing individual district trends (gray lines) and state average trends (colored lines)."}
region_temp_smrs
```

```{r rape_all_m, fig.width=8.5, fig.height=10, fig.cap="Supplementary Figure 2: Comparison of spatial patterns in relative risks (RR) and posterior probabilities (PP) across three models for rape incidents: spatial-only model (top), spatio-temporal model without interaction (middle), and spatio-temporal model with Type I interaction (bottom)."}
# Plot unsmoothed data
rape_all_m
```

```{r dowry_all_m, fig.width=8.5, fig.height=10, fig.cap="Supplementary Figure 3: Comparison of spatial patterns in relative risks (RR) and posterior probabilities (PP) across three models for dowry deaths: spatial-only model (top), spatio-temporal model without interaction (middle), and spatio-temporal model with Type I interaction (bottom)."}
# Plot unsmoothed data
dowry_all_m
```

```{r int_map, fig.width=8.5, fig.height=10, fig.cap="Supplementary Figure 4: Space-time interaction effects for rape incidents (top) and dowry deaths (bottom) in Uttar Pradesh districts from 2001 to 2014, highlighting areas with unusual local temporal trends."}
int_map
```

# References
